\begin{resume}

  \section*{发表的学术论文} % 发表的和录用的合在一起

  \begin{enumerate}[{[}1{]}]
  \addtolength{\itemsep}{-.36\baselineskip}%缩小条目之间的间距，下面类似
  
  \item Zhunchen Luo, Miles Osborne, Sasa Petrovic and Ting Wang. Improving Twitter Retrieval by Exploiting Structural Information. In {\it Proceedings of the Twenty-Sixth AAAI Conference on Artificial Intelligence} (\textbf{AAAI 2012}), Toronto, Canada, July 2012. \textbf{（CCF A类会议，人工智能领域顶级会议）}
 
 \item Zhunchen Luo, Miles Osborne, Jintao Tang and Ting Wang. Who Will Retweet Me? Finding Retweeters in Twitter. In {\it Proceedings of the Thirty-Sixth International ACM SIGIR Conference on Research and Development in Information Retrieval} (\textbf{SIGIR 2013}), Dublin, Ireland, July 2013.\textbf{（CCF A类会议，信息检索领域顶级会议，获得会议旅行奖金1300美元）}

\item Zhunchen Luo, Miles Osborne and Ting Wang. Opinion Retrieval in Twitter. In {\it Proceedings of the Sixth International AAAI Conference on Weblogs and Social Media} (\textbf{AAAI-ICWSM 2012}), Dublin, Ireland, June 2012. \textbf{（社交媒体领域顶级会议，获得会议旅行奖金300美元）}

 \item  Zhunchen Luo, Miles Osborne and Ting Wang. An Effective Approach to Tweets Opinion Retrieval. To appear in \textbf{World Wide Web Journal}. （\textbf{SCI期刊，影响因子1.196}）
 
 \item Zhunchen Luo, Jintao Tang and Ting Wang. Propagated Opinion Retrieval in Twitter. In {\it Proceedings of the Fourteenth International Conference on Web Information System Engineering} (\textbf{WISE 2013}), Nanjing, China, October 2013.\textbf{（CCF C类会议，信息检索与数据挖掘领域重要会议）}

\item Zhunchen Luo, Jintao Tang and Ting Wang. Improving Keyphrase Extraction from Web News by Exploiting Comments Information. In {\it Proceedings of the Fifteenth International Asia-Pacific Web Conference} (\textbf{APWeb 2013}), Sydney, Australia, April 2013. \textbf{（CCF C类会议，信息检索与数据挖掘领域重要会议）}

\item Zhunchen Luo, Lan Rao, Chengsen Ru and Ting Wang. Finding High-Quality Posts from Microblogging Conversations. In {\it the Eighth International Conference on Modeling Decisions for Artificial Intelligence (\textbf{MDAI 2011})}, Changsha, China, July, 2011.

\item 罗准辰，王挺. 基于分离模型的中文关键词提取算法研究. 中文信息学报, 2009，23 (01): 63-70.

\item 罗准辰，王挺. 搜索词同现网络研究. 第六届全国信息检索学术会议(\textbf{CCIR 2010})，镜泊湖，2010年8月.

  \end{enumerate}

\end{resume}
